04_DESeq_sexes

### 2019-08-28 analysis/04_DESeq2_sexes.Rmd
now with LDA functions in functions

```{r LDA}
LDAanalysis <- function(mydds, mycolData, mysubtitle){
  # data prep
  vsd <- vst(mydds, blind=FALSE) # variance stabilized 
  tvsd <- as.data.frame(t(assay(vsd)))   # get vsd counts and transform

  #tvsd <- tvsd[, sample(ncol(tvsd), 20)] # select 20 random genes
  tvsd$V1 <- row.names(tvsd)             # prep for joining

  mydata <- left_join(tvsd, mycolData, by = "V1")  # merge counts and variables
  mydata$V1 <- NULL
  mydata$sextissue <- NULL

  # Split the data into training (80%) and test set (20%)
  set.seed(175)

  training.samples <- mydata$treatment %>%
    createDataPartition(p = 0.8, list = FALSE)
  train.data <- mydata[training.samples, ]
  test.data <- mydata[-training.samples, ]

  # Normalize the data. Categorical variables are automatically ignored.
  # Estimate preprocessing parameters
  preproc.param <- train.data %>% 
    preProcess(method = c("center", "scale"))

  # Transform the data using the estimated parameters
  train.transformed <- preproc.param %>% predict(train.data)
  test.transformed <- preproc.param %>% predict(test.data)

  # LDA analysis
  # Fit the model
  model <- lda(hypothesis~ ., data = train.transformed)
  # Make predictions
  predictions <- model %>% predict(test.transformed)
  # Model accuracy
  mean(predictions$class==test.transformed$hypothesis)

  # results
  print(model)

  # make predictions
  predictions <- model %>% predict(test.transformed)

  # Predicted classes
  head(predictions$class, 6)
  # Predicted probabilities of class memebership.
  head(predictions$posterior, 6) 
  # Linear discriminants
  head(predictions$x, 3)

  # LDA plot using ggplot2 
  lda.data <- cbind(train.transformed, predict(model)$x)
  p <- ggplot(data = lda.data, aes(LD1, LD2, color = hypothesis, shape = sex)) +
    geom_point() +
    labs(subtitle = mysubtitle)
  plot(p)

  # model accuracy 
  print(mean(predictions$class==test.transformed$hypothesis))
}

colDataHyp <- subsetcolData2(c.colData, c("female_hypothalamus", "male_hypothalamus"))
colDataPit <- subsetcolData2(c.colData, c("female_pituitary", "male_pituitary"))
colDataGon <- subsetcolData2(c.colData, c("female_gonad", "male_gonad"))

LDAanalysis(dds.hypothalamus, colDataHyp, "Hypothalamus")
LDAanalysis(dds.pituitary, colDataPit, "Pituitary")
LDAanalysis(dds.gonad, colDataGon, "Gonad")
```

```{r LDA2}

LDAanalysis2 <- function(mydds, mycolData, mysubtitle){
  # data prep
  vsd <- vst(mydds, blind=FALSE) # variance stabilized 
  tvsd <- as.data.frame(t(assay(vsd)))   # get vsd counts and transform

  #tvsd <- tvsd[, sample(ncol(tvsd), 20)] # select 20 random genes
  tvsd$V1 <- row.names(tvsd)             # prep for joining

  mydata <- left_join(tvsd, mycolData, by = "V1")  # merge counts and variables
  mydata$V1 <- NULL
  mydata$sextissue <- NULL
  mydata$hypothesis <- NULL

  # Split the data into training (80%) and test set (20%)
  set.seed(175)

  training.samples <- mydata$treatment %>%
    createDataPartition(p = 0.8, list = FALSE)
  train.data <- mydata[training.samples, ]
  test.data <- mydata[-training.samples, ]

  # Normalize the data. Categorical variables are automatically ignored.
  # Estimate preprocessing parameters
  preproc.param <- train.data %>% 
    preProcess(method = c("center", "scale"))

  # Transform the data using the estimated parameters
  train.transformed <- preproc.param %>% predict(train.data)
  test.transformed <- preproc.param %>% predict(test.data)

  # LDA analysis
  # Fit the model
  model <- lda(treatment~ ., data = train.transformed)
  # Make predictions
  predictions <- model %>% predict(test.transformed)
  # Model accuracy
  mean(predictions$class==test.transformed$treatment)

  # results
  print(model)

  # make predictions
  predictions <- model %>% predict(test.transformed)

  # Predicted classes
  head(predictions$class, 6)
  # Predicted probabilities of class memebership.
  head(predictions$posterior, 6) 
  # Linear discriminants
  head(predictions$x, 3)

  # LDA plot using ggplot2 
  lda.data <- cbind(train.transformed, predict(model)$x)
  p <- ggplot(data = lda.data, aes(LD1, LD2, color = treatment, shape = sex)) +
    geom_point() +
    labs(subtitle = mysubtitle)
  plot(p)

  # model accuracy 
  print(mean(predictions$class==test.transformed$hypothesis))
}

LDAanalysis2(dds.hypothalamus, colDataHyp, "Hypothalamus")
LDAanalysis2(dds.pituitary, colDataPit, "Pituitary")
LDAanalysis2(dds.gonad, colDataGon, "Gonad")

```

>>>>

```{r LDA}
# prep col data for LDA
colDataHyp <- subsetcolData2(c.colData, c("female_hypothalamus", "male_hypothalamus"))
colDataPit <- subsetcolData2(c.colData, c("female_pituitary", "male_pituitary"))
colDataGon <- subsetcolData2(c.colData, c("female_gonad", "male_gonad"))

# LDA for treatment (9 groups)
LDA.hyp1 <- LDAdata.treatment(vsd.hyp, colDataHyp)
LDA.pit1 <- LDAdata.treatment(vsd.pit, colDataPit)
LDA.gon1 <- LDAdata.treatment(vsd.gon, colDataGon)

LDAplot.treatment(LDA.hyp1 , "Hypothalamus")
LDAplot.treatment(LDA.pit1 ,  "Pituitary")
LDAplot.treatment(LDA.gon1 , "Gonad")

# LDA for hypothesis (3 groups)

LDA.hyp1 <- LDAdata.hypothesis(vsd.hyp, colDataHyp)
LDA.pit1 <- LDAdata.hypothesis(vsd.pit, colDataPit)
LDA.gon1 <- LDAdata.hypothesis(vsd.gon, colDataGon)

LDAplot.hypothesis(LDA.hyp1 , "Hypothalamus")
LDAplot.hypothesis(LDA.pit1 ,  "Pituitary")
LDAplot.hypothesis(LDA.gon1 , "Gonad")
```
